Using reference models in feature selection problems \vspace{.1in}

### 2019-11-08 paper/ref_approach.tex

\usepackage{amssymb,amsmath,amsthm}
\usepackage{newtxtext}
\usepackage{newtxmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={The reference model approach in feature selection problems},
            pdfauthor={Federico Pavone, Juho Piironen, Paul-Christian B\"{u}rkner, Aki Vehtari},
            pdfkeywords={keywords},
            pdfborder={0 0 0},
            breaklinks=true,
            colorlinks=true,
            linkcolor=black,
            citecolor=RoyalBlue,
            filecolor=black,
            urlcolor=RoyalBlue,
          }
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=american]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[variant=american]{english}
\fi
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{placeins}
\usepackage{graphicx,grffile,subcaption}
\textwidth 6.0 true in
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{3}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

When the variable selection is repeated with different simulated data
sets, there is some variability in which variables are selected. We
measure the stability of variable selection by computing the entropy
of observed distribution over different models. The smallest entropy
would be obtained if the approach selectes always the same set of
variables, and the largest entropy would be observed if the approach
would always select different set of variables.
Figure \ref{fig:entropy} reports the entropy scores for the different
methods compared.  
The use of a reference model
improves slightly the stability of steplm in feature selection. The
projection predictive approach results to be far more stable than all
other approaches. This is likely due to projpred being based on better
decision theoretical formulation which 1) takes into account the full
predictive distribution and not just point estimate and 2) projects
the reference model posterior to the submodel instead of using a
simple refit of submodels.

@@

We can use the projection predictive approach for complete feature
selection by using it iteratively.  Applying the straightforward
implementation of projpred, we are able to select a minimal subset of
features, which yield to a model with a predictive performance
comparable to the full model one.  The iterative projection repeats
the projpred selection for different iterations, at each time
excluding from the search the features selected in the previous
iterations. At each iteration the selected submodel size corresponds to the
one having a predictive performance close enough to the baseline
model, which in this iterative version is the submodel with the
highest predictive score explored at the current iteration. The
stopping rule that we adopted is the following:
\begin{equation} 
\label{eq:rule_of_thumb}
\text{min} \{i\in \{0,\ldots,p\}: \text{P}(\text{elpd}_{i}-\text{elpd}_{\text{best}}>0)\geq \alpha \},
\end{equation}
where $i$ indexes the submodel size and ``best'' stays for the best predictive explored submodel
at the considered iteration. The algorithm terminates when the empty model (only intercept) satisfies
the stopping rule.
The choice of the hyperparameter $\alpha$ can be non-trivial, we have 
observed sensitivity of the selection to such a choice mainly when using cross-validation with a small
number of observations or not very predictive features. 
In our experiments we chose the default value used in the \texttt{projpred} 
\texttt{R}-package that corresponds to $\alpha=0.16$.
\hl{NOTE: if necessary I can add a pseudo-code illustration of the iterative projection algorithm, as 
I did in the thesis}
TODO: add the pseudo code here

@@

Figure \ref{fig:sensitivity_vs_fdr_iterated} shows the average sensitivity on
the vertical axis and the average false discovery rate on the
horizontal axis based on 100 data simulations for the different
combinations of $n$ and $\rho$. The best selection performance is on the
top-left corner of each plot, as it implies the lowest false discovery
rate and the highest sensitivity. We see that regardless of the
applied selection method, the use of a reference model improves the
selection performance, as it reduces the false discovery rate (shifting
to the left) or increases the sensitivity (shifting upwards). In
accordance with what can be expected, the larger data set size ($n$) and the higher the true correlations ($\rho$), the easier the
selection is. Thus, for easier selection scenarios, the benefits of
the reference model are smaller since the raw data already provide
enough information to identify the relevant covariates. The iterative
projection predictive approach has good false discovery rate in all cases, and the sensitivity is good except when the number of observations and the correlation level
are small. Tuning $\alpha$ might lead to improved selection, yet
the main source of poor performance could well be the miscalibration of the uncertainty of the predictive
utility score for misspecified models with cross-validation \hl{(maybe add some references?)}.

>>>>

\usepackage{amssymb,amsmath,amsthm}
\usepackage{newtxtext}
\usepackage{newtxmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\usepackage{algorithm2e}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={The reference model approach in feature selection problems},
            pdfauthor={Federico Pavone, Juho Piironen, Paul-Christian B\"{u}rkner, Aki Vehtari},
            pdfkeywords={keywords},
            pdfborder={0 0 0},
            breaklinks=true,
            colorlinks=true,
            linkcolor=black,
            citecolor=RoyalBlue,
            filecolor=black,
            urlcolor=RoyalBlue,
          }
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=american]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[variant=american]{english}
\fi
\usepackage{natbib}
\bibliographystyle{plainnat}
\usepackage{placeins}
\usepackage{graphicx,grffile,subcaption}
\textwidth 6.0 true in
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{3}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

When the variable selection is repeated with different simulated data
sets, there is some variability in which variables are selected. We
measure the stability of variable selection by computing the entropy
of observed distribution over different models. The smallest entropy
would be obtained if the approach selectes always the same set of
variables, and the largest entropy would be observed if the approach
would always select different set of variables.
Figure \ref{fig:entropy} reports the entropy scores for the different
methods compared. Lower entropy corresponds to a more stable
selection. Highly correlated predictive features may happen to be
alternately selected, making the measure of the stability of the
selection a non-trivial task. Entropy can not distinguish the interchangeability 
due to correlation from instability, thus such a measure should be 
consider as relative indicator and not as an absolute measure of stability.
\hl{TODO: The above sentence should be clearer now}
The use of a reference model
improves slightly the stability of steplm in feature selection. The
projection predictive approach results to be far more stable than all
other approaches. This is likely due to projpred being based on better
decision theoretical formulation which 1) takes into account the full
predictive distribution and not just point estimate and 2) projects
the reference model posterior to the submodel instead of using a
simple refit of submodels.

@@

We can use the projection predictive approach for complete feature
selection by using it iteratively.  Applying the straightforward
implementation of projpred, we are able to select a minimal subset of
features, which yield to a model with a predictive performance
comparable to the full model one.  The iterative projection repeats
the projpred selection for different iterations, at each time
excluding from the search the features selected in the previous
iterations. At each iteration the selected submodel size corresponds to the
one having a predictive performance close enough to the baseline
model, which in this iterative version is the submodel with the
highest predictive score explored at the current iteration. The
stopping rule that we adopted is the following:
\begin{equation} 
\label{eq:rule_of_thumb}
\text{min} \{i\in \{0,\ldots,p\}: \text{P}(\text{elpd}_{i}-\text{elpd}_{\text{best}}>0)\geq \alpha \},
\end{equation}
where $i$ indexes the submodel size and ``best'' stays for the best predictive explored submodel
at the considered iteration. The algorithm terminates when the empty model (only intercept) satisfies
the stopping rule (see Algorithm \ref{alg:iterated_proj}).
The choice of the hyperparameter $\alpha$ can be non-trivial, we have 
observed sensitivity of the selection to such a choice mainly when using cross-validation with a small
number of observations or not very predictive features. 
In our experiments we chose the default value used in the \texttt{projpred} 
\texttt{R}-package that corresponds to $\alpha=0.16$.

\begin{center}
\begin{algorithm}[H]
\SetAlgoLined
\KwResult{$R:=\{\text{selected features}\}$}
 $F:=\{\text{set of features}\}$ \;
 $R=\{\emptyset\}$ \;
 Fit reference model\;
 \While{$F\neq\{\emptyset\}$}{
   projection.HeuristicSearch \;
   projection.elpdEstimate() \;
%:
   $S=\text{min}\{\text{sub}: P(\text{elpd}_{\text{sub}}-\text{elpd}_{\text{base}}>0)\geq\alpha)$\} \;

  \eIf{$S=\{\emptyset\}$}{
   break \;
   }{
   $R=R\cup S$\;
   $F=F\setminus S$\;
  }
 }
 \caption{Automated iterative projections}
 \label{alg:iterated_proj}
\end{algorithm}
\end{center}

@@

Figure \ref{fig:sensitivity_vs_fdr_iterated} shows the average sensitivity on
the vertical axis and the average false discovery rate on the
horizontal axis based on 100 data simulations for the different
combinations of $n$ and $\rho$. The best selection performance is on the
top-left corner of each plot, as it implies the lowest false discovery
rate and the highest sensitivity. We see that regardless of the
applied selection method, the use of a reference model improves the
selection performance, as it reduces the false discovery rate (shifting
to the left) or increases the sensitivity (shifting upwards). In
accordance with what can be expected, the larger data set size ($n$) and the higher the true correlations ($\rho$), the easier the
selection is. Thus, for easier selection scenarios, the benefits of
the reference model are smaller since the raw data already provide
enough information to identify the relevant covariates. The iterative
projection predictive approach has good false discovery rate in all cases, and the sensitivity is good except when the number of observations and the correlation level
are small. Tuning $\alpha$ might lead to improved selection, yet
the main source of poor performance could well be the miscalibration of the uncertainty of the predictive
utility score for misspecified models with cross-validation \hl{(maybe add some references? The last sentence 
refers to the problem when comparing misspecified models using CV. Tuomas did a talk about this problem.
Maybe Aki can check the content and add some references - not sure Tuomas' work is already available)}.
